EIP CRE transmission network manuscript figures and data

### 2023-06-15 manuscript-figures.Rmd
Updates based on review responses

```{r}
# load in data
# facility and ST
facil_st_orig <- read_csv('data/isolate_data.csv') %>% 
  mutate(cult_year = gsub('-.*|.*/', '', Cult_Date))
# to make analyses more consistent across states
facil_st <- facil_st_orig %>% 
  filter(cult_year >= 2016)
# public isolate data
metadat_complete <- read_csv('data/public_isolate_metadata.csv') 
# public isolate trees
tree_list <- read_rds('data/tree_list.rds')
# pairwise snv distances
kp_dists <- read_rds('data/kp_dists.rds')
ec_dists <- read_rds('data/eh_dists.rds')
# phylogenetic trees
tr_kp <- read.tree('data/kp_tr.tree')
tr_ec <- read.tree('data/eh_tr.tree')
# clinical factors
pt_feats_ct <- read_csv('data/pt_feats_ct.csv')
pt_feats_mn <- read_csv('data/pt_feats_mn.csv')
pt_feats_tn <- read_csv('data/pt_feats_tn.csv')
# shared facility exposures
pr_hosp_pairs_ct <- read_csv('data/pr_hosp_pairs_ct.csv')
pr_hosp_pairs_mn_orig <- read_csv('data/pr_hosp_pairs_mn.csv')
pr_hosp_pairs_mn <- pr_hosp_pairs_mn_orig
pr_hosp_pairs_tn <- read_csv('data/pr_hosp_pairs_tn.csv')
# patient transfer networks
ct_trans_net_sub <- read_csv('data/trans_net_ct.csv')
mn_trans_net_sub <- read_csv('data/trans_net_mn.csv')
tn_trans_net_sub <- read_csv('data/trans_net_tn.csv')
# facility types
facil_types <- read_csv('data/facil_types.csv')
# facility counts
facil_type_counts <- read_csv('data/facil_type_counts.csv')
# prior tn hospitals
prior_hosps_tn <- read_csv('data/pr_hosps_tn.csv')
# tn facil clutser
tn_facil_cluster <- read_csv('data/tn_facil_cluster.csv')
```

@@

@@

```{r}
facil_st %>% 
  group_by(State) %>% 
  summarize(n_facil = n_distinct(Facility_ID))
```

@@

### Isolate counts by state

```{r}
state_dates <- facil_st %>% mutate(Cult_Date = gsub('-.*|.*/','', Cult_Date)) %>% 
  filter(!is.na(Cult_Date)) %>% 
  group_by(State, Cult_Date) %>% summarize() %>% 
  group_by(State) %>% summarize(min_cult_date = min(Cult_Date),
                                max_cult_date = max(Cult_Date))

facil_st %>% 
  mutate(State = factor(State, levels = c('MN','TN','CT'))) %>% 
  group_by(State) %>% 
  count() %>% full_join(state_dates, .) %>% 
  suppressMessages()
```

@@

```{r}
fs1a <- facil_st %>% 
  mutate(ST = ifelse(ST %in% c('E. hormaechei ST114','E. hormaechei ST171', 
                                 'K. pneumoniae ST258', 'K. pneumoniae ST307'), ST, 
                       'Other'),
           ST = factor(ST, levels = rev(c('E. hormaechei ST171', 'E. hormaechei ST114', 
                                      'K. pneumoniae ST258', 'K. pneumoniae ST307', 
                                      'Other')))) %>% 
  group_by(State, ST, cult_year) %>% 
  filter(!is.na(cult_year)) %>% 
  summarize(count = n()) %>% 
  ggplot(aes(x = cult_year, y = count, fill = ST)) +
  facet_grid(~State, scales = 'free', space = 'free') +
  geom_col() +
  scale_fill_manual(values = st_cols, 
                    breaks = c('E. hormaechei ST171', 'E. hormaechei ST114',
                               'K. pneumoniae ST258', 'K. pneumoniae ST307',
                               'Other'), 
                    labels = c('KPC-Eh ST171', 'KPC-Eh ST114',
                               'KPC-Kp ST258', 'KPC-Kp ST307',
                               'Other')) +
  theme(legend.text = element_markdown()) +
  labs(x = 'Culture year', y = 'Number of isolates')
```

@@

```{r, fig.width=5, fig.height=5}
fs1 <- plot_grid(fs1a, fs1b + theme(legend.position = 'none'), ncol = 1, labels = 'AUTO')

ggsave(plot = fs1, filename = 'figures/FigS1.png', width = 5, height = 5)

@@

```{r, fig.width=10, fig.height=8}
fs2 <- plot_grid(fs2ab, fs2c, labels = c('', 'C'), label_x = -0.1)
ggsave('figures/FigS2.png', width = 10, height = 8)
fs2
```

```{r}
kpc_bin <- metadat_complete %>% select(isolate_no, kpc) %>% drop_na() 
mn_bin <- metadat_complete %>% select(isolate_no, mn_bin) %>% drop_na()
tn_bin <- metadat_complete %>% select(isolate_no, tn_bin) %>% drop_na()
ct_bin <- metadat_complete %>% select(isolate_no, ct_bin) %>% drop_na()

@@

keepers <- facil_st %>% 
  filter(!(duplicated(Patient_ID) & duplicated(ST))) %>% 
  pull(Sample_ID)

transitions <- lapply(1:length(tree_list), function(x){

@@

@@

```{r}
transitions <- transitions %>% 
  mutate(ST = ifelse(grepl('ST258', ST_orig), 'K. pneumoniae ST258', as.character(ST))) 
clust_cols <- st_cols[as.character(transitions$ST)]
names(clust_cols) <- paste0(transitions$state, transitions$ST, transitions$cluster)

@@

```{r, fig.width=7, fig.height=4}
f1 <- transitions %>% 
  ggplot(aes(x = class2, 
             fill = factor(paste0(state, ST, cluster), 
                           levels = unique(paste0(state, ST, cluster))))) + 
  geom_bar(color = 'white', position = position_stack(reverse = TRUE)) + 
  scale_fill_manual(values = clust_cols,
                    breaks = rev(names(clust_cols)[!duplicated(clust_cols)]),
                    labels = c('KPC-Eh ST171', 'KPC-Eh ST114',
                               'KPC-Kp ST258', 'KPC-Kp ST307',
                               'Other'),
                    drop = FALSE) +
  facet_grid(~state, scales = 'free', space = 'free') +
  labs(fill = 'ST cluster', x = 'Evidence that KPC-harboring strain was imported', y = 'Number of isolates') +
  theme(legend.text = ggtext::element_markdown())

ggsave(plot = f1, filename = 'figures/Fig1.png', width = 7, height = 4)

f1
```

@@

ggsave(plot = fs3, filename = 'figures/FigS3.png', width = 7.5, height = 10.5)

@@

```{r}
min_dat %>% 
  filter(!is.na(`Inter-facility pair`) & !is.na(`Intra-facility pair`)) %>% 
  group_by(ST_orig, state1) %>% 
  summarize(n = n(), 
            median_intra = median(`Intra-facility pair`),
            median_inter = median(`Inter-facility pair`),
            wilcox_p = wilcox.test(`Intra-facility pair`, `Inter-facility pair`, 
                                   paired = TRUE)$p.value) %>% 
  arrange(-n)
```

@@

```{r}
bind_rows(pr_hosp_pairs_ct, pr_hosp_pairs_mn %>% filter(!is.na(state1)), pr_hosp_pairs_tn) %>%
  left_join(snv_dat %>% filter(loc1 != loc2)) %>% 
  mutate(st1 = ifelse(st1 %in% c("E. hormaechei ST114", "E. hormaechei ST171", "K. pneumoniae ST258", "K. pneumoniae ST307"), st1, 'Other')) %>% 
  left_join(pr_hops_pvals) %>% 
  filter(state1 == 'MN' & st1 == 'E. hormaechei ST171' |
           state1 == 'TN' & st1 %in% c('K. pneumoniae ST258', 'K. pneumoniae ST307', 'E. hormaechei ST114')) %>% 
  select(state1, ST, n_shared, n_not_shared) %>% 
  unique()
```

@@

ggsave(plot = f2, filename = 'figures/Fig2.png', width = 7, height = 7)

@@

```{r, fig.width=7, fig.height=7}
fs4 <- plot_grid(fs4a, fs4b, ncol = 1, labels = 'AUTO')
ggsave(plot = fs4, filename = 'figures/FigS4.png', width = 7, height = 7)
fs4
```

@@

```{r}
mn_snv_dat_orig %>% 
  filter(ST %in% c('ST171') & pairwise_dist <= 10 &
           !is.na(loc1) & !is.na(loc2) & loc1 != loc2) %>% 
  mutate(datediff_bin = cut(as.numeric(datediff), breaks = c(-1, 365*1:7),
                            labels = c('0-1 year', paste0(1:6, '-', 2:7, ' years'))),
         under_thresh = factor(under_thresh, 
                               levels = rev(unique(levels(under_thresh)))),
         loc1 = ifelse(loc1 %in% c('F38'), loc1, 'Other'),
         loc2 = ifelse(loc2 %in% c('F38'), loc2, 'Other')
         # loc1 = ifelse(loc1 %in% c('F38', 'F55', 'F34', 'F46'), loc1, 'Other'),
         # loc2 = ifelse(loc2 %in% c('F38', 'F55', 'F34', 'F46'), loc2, 'Other')
         ) %>% 
  rowwise() %>% 
  mutate(loc1_new = paste(sort(c(loc1, loc2)))[1],
         loc2_new = paste(sort(c(loc1, loc2)))[2]) %>% 
  ungroup() %>%
  mutate(f38 = loc1_new == 'F38' | loc2_new == 'F38') %>% 
  group_by(f38) %>% 
  summarize(count = n()) %>% 
  ungroup() %>% 
  mutate(tot = sum(count),
         pct = paste0(round(count/tot*100), '%')) 
```

```{r}
mn_snv_dat_orig %>% 
  filter(ST %in% c('ST171') & pairwise_dist <= 10 &
           !is.na(loc1) & !is.na(loc2) & loc1 == loc2) %>% 
  mutate(f38 = loc1 == 'F38' | loc2 == 'F38') %>% 
  group_by(f38) %>% 
  summarize(count = n()) %>% 
  ungroup() %>% 
  mutate(tot = sum(count),
         pct = paste0(round(count/tot*100), '%')) 
```

```{r}
f3d <- mn_snv_dat_orig %>% 
  filter(ST %in% c('ST171') & pairwise_dist <= 10 &
           !is.na(loc1) & !is.na(loc2) & loc1 != loc2) %>% 
  mutate(datediff_bin = cut(as.numeric(datediff), breaks = c(-1, 365*1:7),
                            labels = c('0-1 year', paste0(1:6, '-', 2:7, ' years'))),
         under_thresh = factor(under_thresh, 
                               levels = rev(unique(levels(under_thresh)))),
         loc1 = ifelse(loc1 %in% c('F38', 'F55', 'F34', 'F46'), loc1, 'Other'),
         loc2 = ifelse(loc2 %in% c('F38', 'F55', 'F34', 'F46'), loc2, 'Other')
         ) %>% 
  rowwise() %>% 
  mutate(loc1_new = paste(sort(c(loc1, loc2)))[1],
         loc2_new = paste(sort(c(loc1, loc2)))[2]) %>% 
  ungroup() %>%
  group_by(loc1_new, loc2_new) %>% 
  summarize(count = n()) %>% 
  ungroup() %>% 
  mutate(tot = sum(count),
         pct = paste0(round(count/tot*100), '%')) %>% 
  ggplot(aes(x = loc1_new, y = loc2_new, fill = count, label = pct)) + 
  geom_tile() +
  geom_text(col = 'white') +
  labs(x = 'Location of isolate 1', y = 'Location of isolate 2', 
       fill = 'Number of pairs\nwithin 10 SNVs')
```

@@

ggsave(filename = 'figures/Fig4.png', plot = f4, width = 12, height = 5)
f4
```

>>>>

```{r}
# load in data
# facility and ST
facil_st_orig <- read_csv('data/isolate_data.csv') %>% 
  mutate(cult_year = gsub('-.*|.*/', '', Cult_Date))
# to make analyses more consistent across states
facil_st <- facil_st_orig %>% 
  filter(cult_year >= 2016)
# public isolate data
metadat_complete <- read_csv('data/public_isolate_metadata.csv') 
# public isolate trees
tree_list <- read_rds('data/tree_list.rds')
# pairwise snv distances
kp_dists <- read_rds('data/kp_dists.rds')
ec_dists <- read_rds('data/eh_dists.rds')
# phylogenetic trees
tr_kp <- read.tree('data/kp_tr.tree')
tr_ec <- read.tree('data/eh_tr.tree')
# plasmid clusters
mob_plasmids <- read_rds('data/mobtyper/CRE_EIP_mobtyper_mash_nn_plasmid_matrix_050923.RDS')
mob_clusters <- read_rds('data/mobtyper/CRE_EIP_mobtyper_plasmid_cluster_matrix_050923.RDS')
mob_results <- read_rds('data/mobtyper/CRE_EIP_mobtyper_results_clean_KPC_050923.RDS')
# mob_contigs <- read_rds('data/mobtyper/CRE_EIP_mobtyper_contigs_clean_KPC_050923.RDS') 
# clinical factors
pt_feats_ct <- read_csv('data/pt_feats_ct.csv')
pt_feats_mn <- read_csv('data/pt_feats_mn.csv')
pt_feats_tn <- read_csv('data/pt_feats_tn.csv')
# shared facility exposures
pr_hosp_pairs_ct <- read_csv('data/pr_hosp_pairs_ct.csv')
pr_hosp_pairs_mn_orig <- read_csv('data/pr_hosp_pairs_mn.csv')
pr_hosp_pairs_mn <- pr_hosp_pairs_mn_orig
pr_hosp_pairs_tn <- read_csv('data/pr_hosp_pairs_tn.csv')
# patient transfer networks
ct_trans_net_sub <- read_csv('data/trans_net_ct.csv')
mn_trans_net_sub <- read_csv('data/trans_net_mn.csv')
tn_trans_net_sub <- read_csv('data/trans_net_tn.csv')
# facility types
facil_types <- read_csv('data/facil_types.csv')
# facility counts
facil_type_counts <- read_csv('data/facil_type_counts.csv')
# prior tn hospitals
prior_hosps_tn <- read_csv('data/pr_hosps_tn.csv')
# tn facil clutser
tn_facil_cluster <- read_csv('data/tn_facil_cluster.csv')
```

@@

### Number of isolates

```{r}
nrow(facil_st)
```
### Number of patients

```{r}
n_distinct(facil_st$Patient_ID)
```

### Number of facilities

```{r}
n_distinct(facil_st$Facility_ID)
```

@@

```{r}
facil_st %>% 
  group_by(State) %>% 
  summarize(n_isolates = n(),
            n_facil = n_distinct(Facility_ID))
```

@@

@@

```{r}
fs1a <- facil_st_orig %>% 
  mutate(ST = ifelse(ST %in% c('E. hormaechei ST114','E. hormaechei ST171', 
                                 'K. pneumoniae ST258', 'K. pneumoniae ST307'), ST, 
                       'Other'),
           ST = factor(ST, levels = rev(c('E. hormaechei ST171', 'E. hormaechei ST114', 
                                      'K. pneumoniae ST258', 'K. pneumoniae ST307', 
                                      'Other')))) %>% 
  group_by(State, ST, cult_year) %>% 
  filter(!is.na(cult_year)) %>% 
  summarize(count = n()) %>% 
  ggplot(aes(x = cult_year, y = count, fill = ST)) +
  facet_grid(~State, scales = 'free', space = 'free') +
  geom_col() +
  scale_fill_manual(values = st_cols, 
                    breaks = c('E. hormaechei ST171', 'E. hormaechei ST114',
                               'K. pneumoniae ST258', 'K. pneumoniae ST307',
                               'Other'), 
                    labels = c('KPC-Eh ST171', 'KPC-Eh ST114',
                               'KPC-Kp ST258', 'KPC-Kp ST307',
                               'Other')) +
  theme(legend.text = element_markdown()) +
  labs(x = 'Culture year', y = 'Number of isolates')
```

@@

### Number of distinct STs by state and year

```{r}
fs1c <- facil_st_orig %>% 
  group_by(State, cult_year, Species, ST) %>% 
  filter(!is.na(cult_year)) %>% 
  summarize() %>% 
  group_by(State, cult_year, Species) %>% 
  count() %>% 
  mutate(Species = ifelse(grepl('hormaechei', Species), 'KPC-Eh', 'KPC-Kp')) %>% 
  ggplot(aes(x = cult_year, y = n, fill = Species)) +
  facet_grid(~State, scales = 'free', space = 'free') +
  geom_col() +
  scale_fill_manual(values = rev(c('coral2', 'aquamarine2'))) +
  labs(x = 'Year', y = 'Number of distinct STs')
```

```{r, fig.width=5.5, fig.height=7}
fs1 <- plot_grid(fs1a, fs1b + theme(legend.position = 'none'), fs1c, ncol = 1, labels = 'AUTO')

ggsave(plot = fs1, filename = 'figures/FigS1.png', width = 5.5, height = 7)
ggsave(plot = fs1, filename = 'figures/FigS1.tiff', width = 5.5, height = 7)

@@

```{r}
# identifying phylogenetic transition events
kpc_bin <- metadat_complete %>% select(isolate_no, kpc) %>% drop_na() 
mn_bin <- metadat_complete %>% select(isolate_no, mn_bin) %>% drop_na()
tn_bin <- metadat_complete %>% select(isolate_no, tn_bin) %>% drop_na()
ct_bin <- metadat_complete %>% select(isolate_no, ct_bin) %>% drop_na()

@@

keepers <- facil_st %>% 
  filter(!(duplicated(Patient_ID) & duplicated(ST) & duplicated(Facility_ID))) %>% 
  pull(Sample_ID)

# phylogenetic clusters
transitions <- lapply(1:length(tree_list), function(x){

@@

```{r}
# plasmid info 
mob_plasmids_long <- mob_plasmids %>% 
  rownames_to_column(var = 'isolate_no') %>% 
  pivot_longer(!isolate_no, names_to = 'mash_nearest_neighbor', values_to = 'present') %>% 
  left_join(mob_results %>% 
              filter(!grepl('novel', primary_cluster_id)) %>% 
              select(primary_cluster_id, mash_nearest_neighbor) %>% distinct()) %>% 
  left_join(mob_results %>% 
              filter(!grepl('novel', primary_cluster_id)) %>% 
              select(isolate_no, primary_cluster_id, mash_nearest_neighbor, KPC_present)) %>% 
  mutate(state = gsub('_.*', '', isolate_no), .before = 2) %>% 
  group_by(mash_nearest_neighbor) %>% 
  filter(any(sum(KPC_present, na.rm = TRUE) > 0)) %>% 
  left_join(read_csv('data/isolate_data.csv') %>% 
  select(Sample_ID, Species, ST) %>% 
  rename(isolate_no = Sample_ID)) %>% 
  mutate(ST_orig = ST,
         ST = case_when(is.na(ST) ~ 'Other species',
                        ST %in% dominant_sts ~ ST, 
                        TRUE ~ 'Other')) %>% 
  ungroup()

cluster_info <- mob_plasmids_long %>% 
  filter(KPC_present == 1) %>%
  rename(plasmid_cluster = primary_cluster_id) %>% 
  inner_join(transitions %>% 
               mutate(phylo_cluster = paste0(state, ST, cluster, class2)) %>% 
               select(id, state, phylo_cluster, class2) %>% 
               rename(isolate_no = id)) %>% 
  select(isolate_no, state, ST, class2, phylo_cluster, plasmid_cluster) %>% 
  group_by(state, ST, phylo_cluster, plasmid_cluster) %>% 
  mutate(cluster_id = paste0(state, ST, phylo_cluster, plasmid_cluster),
         cluster_size = n()) %>% 
  ungroup()

nodes <- cluster_info %>%
               select(ST, state, class2, phylo_cluster, plasmid_cluster) %>%
               pivot_longer(!c(ST, class2, state), names_to = 'node_type', values_to = 'clust_id') %>%
  group_by(node_type, clust_id) %>% 
  summarize(n_in_clust = n()) %>% 
  ungroup() %>% 
  arrange(clust_id) %>% 
  left_join(cluster_info %>% select(ST, class2, phylo_cluster) %>% rename(clust_id = phylo_cluster) %>% distinct()) %>% 
  mutate(index = as.numeric(factor(clust_id)),
         node_type = ifelse(grepl('phylo', node_type), ST, 'Plasmid'),
         node_import = factor(ifelse(is.na(class2), 'Plasmid', as.character(class2)),
                              levels = c('Yes', 'No', 'Plasmid')))

edges <- cluster_info %>% 
  left_join(nodes %>% rename(phylo_cluster = clust_id) %>% rename(from = index) %>% select(-c(node_type, n_in_clust, ST, class2, node_import))) %>%
  left_join(nodes %>% rename(plasmid_cluster = clust_id) %>% rename(to = index) %>% select(-c(node_type, n_in_clust, ST, class2, node_import))) %>% 
  group_by(from, to, phylo_cluster, plasmid_cluster, state, ST, class2) %>% 
  summarize(n_in_grp = n())

clust_graph <- tbl_graph(directed = FALSE, 
             nodes = nodes,
             edges = edges)
```

```{r}
fs2d <- cluster_info %>%
  group_by(phylo_cluster) %>%
  filter(n() > 1) %>%
  group_by(state, ST, phylo_cluster, plasmid_cluster, cluster_size) %>% 
  summarize() %>% 
  ggplot(aes(x = phylo_cluster, y = cluster_size, group = plasmid_cluster, fill = ST)) +
  facet_grid(~state, scales = 'free', space = 'free') +
  geom_col(col = 'white') +
  scale_fill_manual(values = st_cols,
                    breaks = c('E. hormaechei ST171', 'E. hormaechei ST114',
                               'K. pneumoniae ST258', 'K. pneumoniae ST307',
                               'Other'), 
                    labels = c('KPC-Eh ST171', 'KPC-Eh ST114',
                               'KPC-Kp ST258', 'KPC-Kp ST307',
                               'Other'),
                    drop = FALSE) +
  theme(axis.text.x = element_blank()) +
  labs(x = 'Phylogenetic cluster', y = 'Number of isolates')
```

```{r, fig.width=8, fig.height=10}
fs2 <- plot_grid(plot_grid(fs2ab, fs2c, labels = c('', 'C'), label_x = -0.1),
                  fs2d, ncol = 1, labels = c('', 'D'), rel_heights = c(2, 0.5))
ggsave('figures/FigS2.png', width = 8, height = 10)
ggsave('figures/FigS2.tiff', width = 8, height = 10)
fs2
```

```{r}
cluster_info %>%
  group_by(phylo_cluster) %>%
  filter(n() > 1) %>%
  group_by(state, ST, phylo_cluster, plasmid_cluster, cluster_size) %>% 
  summarize() %>% 
  group_by(state, phylo_cluster) %>% 
  tally() %>% 
  group_by(state, n > 1) %>% 
  tally() %>% 
  # group_by(state) %>%
  # group_by(`n > 1`) %>% 
  ungroup() %>%
  mutate(tot = sum(n))
```

@@

```{r}
# cluster_info <- cluster_info %>% 
  # mutate(ST = ifelse(grepl('ST258', ST_orig), 'K. pneumoniae ST258', as.character(ST))) 
clust_cols <- st_cols[as.character(cluster_info$ST)]
names(clust_cols) <- cluster_info$cluster_id #paste0(cluster_info$state, cluster_info$ST, cluster_info$cluster)

@@

```{r}
set.seed(2)
clust_graph_ct <- clust_graph %>%
  activate(edges) %>%
  filter(state == 'CT') %>% 
  activate(nodes) %>% 
  filter(centrality_degree() > 0) %>% 
  mutate(n_in_clust = centrality_degree(weights = n_in_grp)) %>% 
  ggraph(layout = 'fr') +
    geom_edge_link(aes(alpha = n_in_grp)) +
    geom_node_point(aes(col = node_type, size = n_in_clust)) +
  theme_graph() +
  scale_color_manual(values = c(st_cols, Plasmid = 'grey20'),
                    breaks = c('E. hormaechei ST171', 'E. hormaechei ST114',
                               'K. pneumoniae ST258', 'K. pneumoniae ST307',
                               'Other', 'Plasmid'), 
                    labels = c('KPC-Eh ST171', 'KPC-Eh ST114',
                               'KPC-Kp ST258', 'KPC-Kp ST307',
                               'Other', 'Plasmid'),
                    drop = FALSE) +
  scale_edge_alpha_continuous(limits = c(0, 20), breaks = seq(5, 20, by = 5)) +
  scale_size_continuous(limits = c(0, 50), breaks = c(10, 20, 30), labels = c(10, 20, 30)) +
  labs(col = 'Cluster type', size = 'Number of isolates', edge_alpha = 'Number of\nshared isolates') +
  theme(title = element_text(size = 10))

clust_graph_mn <- clust_graph %>%
  activate(edges) %>%
  filter(state == 'MN') %>% 
  activate(nodes) %>% 
  filter(centrality_degree() > 0) %>% 
  mutate(n_in_clust = centrality_degree(weights = n_in_grp)) %>% 
  ggraph(layout = 'fr') +
    geom_edge_link(aes(alpha = n_in_grp)) +
    geom_node_point(aes(col = node_type, size = n_in_clust)) +
  theme_graph() +
  scale_color_manual(values = c(st_cols, Plasmid = 'grey20'),
                    breaks = c('E. hormaechei ST171', 'E. hormaechei ST114',
                               'K. pneumoniae ST258', 'K. pneumoniae ST307',
                               'Other', 'Plasmid'), 
                    labels = c('KPC-Eh ST171', 'KPC-Eh ST114',
                               'KPC-Kp ST258', 'KPC-Kp ST307',
                               'Other', 'Plasmid'),
                    drop = FALSE) +
  scale_edge_alpha_continuous(limits = c(0, 20), breaks = seq(5, 20, by = 5)) +
  scale_size_continuous(limits = c(0, 50), breaks = c(10, 20, 30), labels = c(10, 20, 30)) +
  labs(col = 'Cluster type', size = 'Number of isolates', edge_alpha = 'Number of\nshared isolates')

clust_graph_tn <- clust_graph %>%
  activate(edges) %>%
  filter(state == 'TN') %>% 
  activate(nodes) %>% 
  filter(centrality_degree() > 0) %>%
  mutate(n_in_clust = centrality_degree(weights = n_in_grp)) %>% 
  ggraph(layout = 'fr') +
    geom_edge_link(aes(alpha = n_in_grp)) +
    geom_node_point(aes(col = node_type, size = n_in_clust)) +
  theme_graph() +
  scale_color_manual(values = c(st_cols, Plasmid = 'grey20'),
                    breaks = c('E. hormaechei ST171', 'E. hormaechei ST114',
                               'K. pneumoniae ST258', 'K. pneumoniae ST307',
                               'Other', 'Plasmid'), 
                    labels = c('KPC-Eh ST171', 'KPC-Eh ST114',
                               'KPC-Kp ST258', 'KPC-Kp ST307',
                               'Other', 'Plasmid'),
                    drop = FALSE) +
  scale_edge_alpha_continuous(limits = c(0, 20), breaks = seq(5, 20, by = 5)) +
  scale_size_continuous(limits = c(0, 50), breaks = c(10, 20, 30), labels = c(10, 20, 30)) +
  labs(col = 'Cluster type', size = 'Number of isolates', edge_alpha = 'Number of\nshared isolates')

f1a <- plot_grid(plot_grid(clust_graph_ct + 
                   theme(legend.position = 'none', 
                         plot.title = element_text(hjust = 0.5, size = 10, face = 'plain')) +
            labs(title = 'CT'), 
          clust_graph_mn + 
            theme(legend.position = 'none', 
                  plot.title = element_text(hjust = 0.5, size = 10, face = 'plain')) +
            labs(title = 'MN'), 
          clust_graph_tn + 
            theme(legend.position = 'none', 
                  plot.title = element_text(hjust = 0.5, size = 10, face = 'plain')) +
            labs(title = 'TN'), 
          ncol = 1), get_legend(clust_graph_mn + theme(legend.position = 'right')), nrow = 1) %>% 
  suppressWarnings()
```

```{r, fig.width=7, fig.height=4}
f1b <- cluster_info %>% 
  arrange(ST, -cluster_size) %>% 
  ggplot(aes(x = class2, 
             fill = factor(cluster_id, #paste0(state, ST, cluster), 
                           levels = unique(cluster_id)))) + #unique(paste0(state, ST, cluster))))) + 
  geom_bar(color = 'white', position = position_stack(reverse = TRUE)) + 
  scale_fill_manual(values = clust_cols,
                    breaks = rev(names(clust_cols)[!duplicated(clust_cols)]),
                    labels = c('KPC-Eh ST171', 'KPC-Eh ST114',
                               'KPC-Kp ST258', 'KPC-Kp ST307',
                               'Other'),
                    drop = FALSE) +
  facet_grid(~state, scales = 'free', space = 'free') +
  labs(fill = 'ST cluster', x = 'Evidence that KPC-harboring strain was imported', y = 'Number of isolates') +
  theme(legend.text = ggtext::element_markdown())
```

```{r, fig.width=12, fig.height=6}
# f1 <- plot_grid(f1a, plot_grid(f1b, NULL, nrow = 1, rel_widths = c(2, 0.5)), 
#                 ncol = 1, rel_heights = c(2, 1), labels = c('A', 'B'))

networks <- plot_grid(plot_grid(clust_graph_ct + 
                   theme(legend.position = 'none', 
                         plot.title = element_text(hjust = 0.5, size = 10, face = 'plain')) +
            labs(title = 'CT'), 
          clust_graph_mn + 
            theme(legend.position = 'none', 
                  plot.title = element_text(hjust = 0.5, size = 10, face = 'plain')) +
            labs(title = 'MN'), nrow = 1), 
          plot_grid(NULL,
          clust_graph_tn + 
            theme(legend.position = 'none', 
                  plot.title = element_text(hjust = 0.5, size = 10, face = 'plain')) +
            labs(title = 'TN'),
          NULL, nrow = 1, rel_widths = c(0.5, 1, 0.5)),
          ncol = 1)

f1 <- plot_grid(networks, 
                get_legend(clust_graph_mn + theme(legend.position = 'right')),
                f1b + theme(legend.position = 'none'), labels = c('A', '', 'B'), 
                nrow = 1, rel_widths = c(2, 0.5, 2)) %>% 
  suppressWarnings()

ggsave(plot = f1, filename = 'figures/Fig1.png', width = 12, height = 6)
ggsave(plot = f1, filename = 'figures/Fig1.tiff', width = 12, height = 6)
```

@@

ggsave(plot = fs3, filename = 'figures/FigS3.png', width = 7.5, height = 10.5)
ggsave(plot = fs3, filename = 'figures/FigS3.tiff', width = 7.5, height = 10.5)

@@

@@

@@

ggsave(plot = f2, filename = 'figures/Fig2.png', width = 7, height = 7)
ggsave(plot = f2, filename = 'figures/Fig2.tiff', width = 7, height = 7)

@@

```{r, fig.width=7, fig.height=7}
fs4 <- plot_grid(fs4a, fs4b, ncol = 1, labels = 'AUTO')
ggsave(plot = fs4, filename = 'figures/FigS4.png', width = 7, height = 7)
ggsave(plot = fs4, filename = 'figures/FigS4.tiff', width = 7, height = 7)
fs4
```

@@

```{r}
mn_snv_dat_orig %>% 
  filter(ST %in% c('ST171') & pairwise_dist <= 5 &
           !is.na(loc1) & !is.na(loc2) & loc1 != loc2) %>% 
  mutate(datediff_bin = cut(as.numeric(datediff), breaks = c(-1, 365*1:7),
                            labels = c('0-1 year', paste0(1:6, '-', 2:7, ' years'))),
         under_thresh = factor(under_thresh, 
                               levels = rev(unique(levels(under_thresh)))),
         loc1 = ifelse(loc1 %in% c('F38'), loc1, 'Other'),
         loc2 = ifelse(loc2 %in% c('F38'), loc2, 'Other')
         # loc1 = ifelse(loc1 %in% c('F38', 'F55', 'F34', 'F46'), loc1, 'Other'),
         # loc2 = ifelse(loc2 %in% c('F38', 'F55', 'F34', 'F46'), loc2, 'Other')
         ) %>% 
  rowwise() %>% 
  mutate(loc1_new = paste(sort(c(loc1, loc2)))[1],
         loc2_new = paste(sort(c(loc1, loc2)))[2]) %>% 
  ungroup() %>%
  mutate(f38 = loc1_new == 'F38' | loc2_new == 'F38') %>% 
  group_by(f38) %>% 
  summarize(count = n()) %>% 
  ungroup() %>% 
  mutate(tot = sum(count),
         pct = paste0(round(count/tot*100), '%')) 
```

```{r}
mn_snv_dat_orig %>% 
  filter(ST %in% c('ST171') & pairwise_dist <= 5 &
           !is.na(loc1) & !is.na(loc2) & loc1 == loc2) %>% 
  mutate(f38 = loc1 == 'F38' | loc2 == 'F38') %>% 
  group_by(f38) %>% 
  summarize(count = n()) %>% 
  ungroup() %>% 
  mutate(tot = sum(count),
         pct = paste0(round(count/tot*100), '%')) 
```

```{r}
f3d <- mn_snv_dat_orig %>% 
  filter(ST %in% c('ST171') & pairwise_dist <= 5 &
           !is.na(loc1) & !is.na(loc2) & loc1 != loc2) %>% 
  mutate(datediff_bin = cut(as.numeric(datediff), breaks = c(-1, 365*1:7),
                            labels = c('0-1 year', paste0(1:6, '-', 2:7, ' years'))),
         under_thresh = factor(under_thresh, 
                               levels = rev(unique(levels(under_thresh))))
         ) %>% 
  rowwise() %>% 
  mutate(f38_loc = any(c(loc1, loc2) == 'F38'),
         loc1_new = ifelse(f38_loc, 
                           c(loc1, loc2)[c(loc1, loc2) == 'F38'],
                           paste(sort(c(loc1, loc2)))[1]),
         loc2_new = ifelse(f38_loc, 
                           c(loc1, loc2)[c(loc1, loc2) != 'F38'],
                           paste(sort(c(loc1, loc2)))[2]),
         loc1_new = ifelse(loc1_new %in% c('F38'), loc1_new, 'Other')) %>% #ifelse(loc2_new %in% c('F38', 'F55', 'F34', 'F46'), loc2_new, 'Other')) %>%
  filter(loc1_new == 'F38') %>% 
  ungroup() %>%
  group_by(loc1_new, loc2_new) %>% 
  summarize(count = n()) %>% 
  ungroup() %>% 
  mutate(tot = sum(count),
         pct = paste0(round(count/tot*100), '%')) %>% 
  arrange(count/tot) %>% 
  mutate(loc2_new = factor(loc2_new, levels = unique(loc2_new))) %>% 
  ggplot(aes(x = loc1_new, y = loc2_new, fill = count, label = pct)) + 
  geom_tile() +
  geom_text(col = 'white', size = 3) +
  labs(x = 'Location of isolate 1', y = 'Location of isolate 2', 
       fill = 'Number of pairs\nwithin 5 SNVs')
```

@@

ggsave(filename = 'figures/Fig4.png', plot = f4, width = 12, height = 5)
ggsave(filename = 'figures/Fig4.tiff', plot = f4, width = 12, height = 5)
f4
```
